The Aquarius Stream Progenitor Was Not A Globular Cluster: \\ Its History is Far More Interesting

### 2012-12-04 2012-aqs/draft-aqs.tex
Added Na-O text

>>>>

% establish the relationship as a true phenomena

Extensive studies looking at stars in globular clusters have revealed variations in light-element abundances, most notably an anti-correlation in sodium and oxygen content \citet{Norris;Da-Costa_1995}. Between generations of stars within a globular cluster, the oxygen abundance is transformed into nitrogen, reducing the overall oxygen abundance. 
This pattern has been repeatedly demonstrated in all well-studied globular clusters, although the strength of the anti-correlation varies between systems. Even though oxygen undergoes internal mixing during its evolution along the giant branch, it is now well established that any observed abundance variations due to such mixing is limited to Li, C, and N. Therefore, the observed anti-correlations in light elements required a different astrophysical explanation.

In order to help explain this phenomena, \citet{Caretta;et-al_2009} defined three components of stars within a population: primordial, intermediate, and extreme. The [Na/Fe] and [O/Fe] abundances for a given star defines which component it belongs to. Primordial (first-generation) stars are those with Na and O content similar to field stars of the same metallicity. For a well-studied system, stars with the lowest cluster [Na/Fe] abundances are characterised as primordial. An intermediate population -- the second-generation of stars -- will have super-solar sodium abundances, and demonstrate a slight depletion in oxygen content. The  abundance exhaustion rate is ultimately dependent on the star-formation history of the cluster environment. The intermediate population is sub-divided based on how much the O and Na abundances depart from the primordial population: stars with [O/Na]$ > -0.9$ dex are classified as belonging to an extreme component of this Na-O pattern because they differ greatly from the primordial cluster abundance. The extreme O-depletion component has only been observed in a few clusters (more massive ones preferentially), indicating an extended star-formation history in more massive clusters.

% discuss the production of Sodium
The increase in Na abundance between generations of stars is a natural consequence of nucleosynthesis in massive ($>8 M_\odot$) stars. Massive stars undergo hydrostatic carbon burning, producing sodium by the dominant $^{12}\mbox{C}(^{12}\mbox{C}, p)^{23}\mbox{Na}$ process and eventually delivering it to the interstellar-medium through SN II events. The total amount of Na production is dependent on the neutron excess of the star, which slowly increases during the star's evolution. Because the eventual SN II explosion is devoid of any significant beta-decay processes, the neutron excess of the exploded material is representative of the pre-explosion abundance. This material eventually condenses to form the next generation of stars, which will have a net increase in their neutron excess with respect to their predecessors. Since the Na-production rate is correlated with the stellar neutron excess, an overall increase in Na is expected. The sodium content also becomes important for production of Ni during the SN II event (see \S\ref{sec:na-ni}) because $^{23}$Na is the only stable isotope produced in significant quantities during the C- (or O-) burning stages.

% discuss the production of Oxygen

% Discuss the depletion of oxygen over time. What are the possible astrophysical explanations for this depletion
The exact astrophysical mechanism for oxygen depletion between generations of stars in globular clusters is still under debate. However, it is clear that this trend has been well-demonstrated by numerous independent observational studies.

Additionally, oxygen is a particularly difficult element to measure. There are only a handful of lines available in an optical spectrum: the forbidden [O \textsc{I}] lines at 630 nm and 636 nm, and the O \textsc{I} triplet lines at 777 nm. The forbidden lines are very weak, and become immeasurable in either hot and/or metal-poor stars ([Fe/H] $\lesssim -1.5$ dex). When they are present, the forbidden lines can be hugely affected by telluric absorption -- depending on the radial velocity of the star. Moreover, the 630 nm line is blended with a Ni \textsc{I} absorption line \citep{Allende-Prieto;et-al_2001}, and hence the region requires a careful synthesis approach. Although the O \textsc{I} triplet lines at 777 nm are stronger than the forbidden lines, they are extremely sensitive to NLTE effects. Oxygen abundances measured using the triplet lines under the assumption of LTE can be $\sim$0.25 dex higher than abundances derived from the forbidden lines \citep{Perez;et-al_2006}.

% Thus if you want to deduce the history of star formation for a group of stars, and deduce whether it has undergone star formation typically seen in globular clusters, you must sample from all components.

It is clear that careful consideration must be given inferring star-formation histories from observed stellar abundances. In addition to the normal care afforded for measuring elemental abundances from high-resolution spectroscopic data, attention must be given to telluric absorption, Ni contamination, and NLTE corrections when deriving oxygen abundances. Furthermore, when characterising the oxygen depletion rate -- and hence the strength of the Na-O anti-correlation for a cluster -- it is vital to obtain observations for stars belonging to all components (where possible). If only a primordial sample of stars in a globular cluster is observed, then their [Na/Fe] and [O/Fe] abundances will be indistinguishable from halo stars of a similar metallicity. In such a scenario, any inferred anti-correlation could equally be explained by intrinsic chemical scatter or observational uncertainties.

% Talk about Wylie de boer measurements

% Wylie de boer claim from their Na-O abundances that...
% Wylie de boer abundances themselves do not demonstrate the classical Na-O anti-correlation. In fact, they have a slight positive relationship. However, this trend is drawn from a sample of 5 stars -- the same as this study. If there is any evidence of a globular cluster from the [Na/Fe], [O/Fe] abundances in the Wylie de Boer study, it is the fact that two of their stars have slightly super-solar Na abundances; which is higher than expected for halo stars of this metallicity.

% How our measurements of Na, O compare to theirs

% Can these abundance differences be explained by a line-list?

% In any case, suppose these stars did originate from a globular cluster. If the entire cluster was well-studied, the primordial component cut-off is described by Caretta as stars which are 4\sigma from the halo abundance at that metallicity. These stars would

In such a scenario, these stars would be classified as a primordial component -- thus being indistinguishable from the halo. It is insufficient
